Existence of Forward Simulations for Particular Data Structures

### 2017-01-12 draftCAV/prelim.tex
Comments on Preliminaries

A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, s_0, \delta)$ over the 
possibly-infinite alphabet $\Sigma$ is a possibly-infinite set $Q$ of states with
initial state $s_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
sequence is denoted by $\epsilon$.
An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ for any $0\leq i\leq j <k$
(in particular $s_i\xrightarrow{\epsilon}s_i$).
The projection $@t| \Gamma$ of a sequence $@t$ is the maximum subsequence of $@t$ over
 alphabet $\Gamma$. This notation is extended to sets of sequences as usual.
A \emph{trace} of $A$ is the projection $\rho | \Sigma$ of an execution $\rho$ of $A$. 
The set of executions, resp., traces, of an LTS $A$ is denoted by $E(A)$, resp., $Tr(A)$.
An LTS is \emph{deterministic} if for any state $s$ and any sequence $@t\in \Sigma^*$, there is at most
one state $s'$ such that $s\xrightarrow{@t}s'$.

@@

\noindent
We fix an arbitrary set $\<Ops>$ of operation identifiers, and for given sets
$\<Methods>$ and $\<Vals>$ of methods and values, we fix the sets
\begin{align*}
  & C = \set{ inv(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }
  \text{, and } \\
  & R = \set{ ret(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }  
\end{align*}
of \emph{call actions} and \emph{return actions}; each call action $inv(m,d,k)$
combines a method $m \in \<Methods>$ and value $d \in \<Vals>$ with an
\emph{operation identifier} $k \in \<Ops>$. Operation identifiers are used to
pair call and return actions. We denote the operation identifier of a
call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
We may omit the second field from a call/return action $a$ for methods that have no inputs or return values 
(e.g., the pop method of a stack).
A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
\emph{well formed} when:
\begin{itemize}

@@

Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. This weakening relation
between histories can be formalized as follows: a history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
if{f} there exists a well-formed history $h_1'$
obtained from $h_1$ by appending return actions, and deleting call actions,
such that:
\begin{quote}

@@

\end{quote}
$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
history set $H$ under weakening.
A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
every library $L$. A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$.
A library $L$ is called atomic if it has a sequential kernel. 
Atomic libraries are often considered as specifications for concurrent objects. 
In practice, libraries can be made atomic by guarding their methods bodies with global lock acquisitions.

@@

Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
is formally captured by \emph{(observational) refinement}. Informally, given two libraries
$L_1$ and $L_2$ implementing the methods of some concurrent object, we
say $L_1$ refines $L_2$ if{f} every computation of every program
using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
Refinement can be proved using \emph{forward/backward simulation relations} which roughly, guarantee that steps
in the concrete library correspond to (sequences of) steps in the abstract library.

\begin{dfn}
Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share $\Gamma$.}. 
%agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. We define the set $A\Sigma = \{inv(m,d,i)| m \in \mathcal{M} \wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\} \cup \{ret(m,d,i)| m \in \mathcal{M}\wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\}$ as abstract transition labels. Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$. 
The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}
The actions in $\Gamma$ are called \emph{observable actions}. Usually, $\Gamma$ consists only of call and return actions. As we show in the next sections, increasing the set of observable actions is useful in simplifying proofs of refinement.
%Linearizability is also a relation between two libraries and it is stricter than refinement. It requires $e'$ in Definition 2 to be a sequential one. A trace $e$ is sequential iff following two conditions hold for its projection to abstract transition labels $e|A\Sigma = e_1, ...e_n$: $(i)$ $e_1 = inv(m,d,k)$ for some $m \in \mathcal{M}, d \in \mathcal{D} and k \in \mathbb{N}$ and $(ii)$ for all $i \in [1,n)$, either $e_i = inv(m,d,k)$ and $e_{i+1} = ret(m,d',k)$ or $e_i = ret(m,d,k)$ and $e_{i+1} = inv(m',d',k')$ for some $m,m' \in \mathcal{M}$, $d,d' \in \mathcal{D}$ and $k,k' \in \mathbb{N}$.

@@

The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. 
%A linearization point 
%corresponds to a transition in that
%execution that occurs between the call and the return action of the operation.
A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
An LP-annotation is called \emph{complete} when every completed operation has a linearization point, i.e., TODO.
An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ when  $\hat{h}$ contains $lin(\_,\_,k)$.

@@

We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic. A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. The existence of backward simulations is equivalent to refinement for arbitrary libraries.
%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.

@@

\begin{dfn}
Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$.
\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$, for each $0\leq i<|@s|$.
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation establishes that $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps. In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action. The rest of the transitions in $L_1$ must be simulated by a possibly empty sequence of transitions of $L_2$ (with arbitrary labels).

>>>>

A \emph{labeled transition system} (LTS) $A=(Q,\Sigma, s_0, \delta)$ over the 
possibly-infinite alphabet $\Sigma$ is a possibly-infinite set $Q$ of states with
initial state $s_0 \in Q$, and a transition relation $\delta \subseteq Q \times @S \times
Q$. The $i$th symbol of a sequence $@t \in @S^*$ is denoted $@t_i$, and the empty
sequence is denoted by $\epsilon$.
An \emph{execution} of $A$ is an alternating sequence of states and transition labels (called also actions)
$\rho = s_0, e_0,s_1\ldots e_{k-1},s_k$ for some $k>0$ such that $\delta(s_i, e_i, s_{i+1})$
for each $i$ such that $0\leq i<k$. We write $s_i\xrightarrow{e_i\ldots e_{j-1}} s_j$ \textcolor{red}{as shorthand for $s_i,e_i,...,s_{j-1},e_{j-1},s_j$ subsequence of $\rho$,} for any $0\leq i\leq j <k$
(in particular $s_i\xrightarrow{\epsilon}s_i$)\textcolor{red}{(I also propose to add subscript to the arrow mentioning the machine name like $s_i \xrightarrow{\alpha}_A s_j$ to avoid confusion between machines in the forward simulation definition. This will require us to change arrows in forward and backward simulation definitions.)}.
The projection $@t| \Gamma$ of a sequence $@t$ is the maximum subsequence of $@t$ over
 alphabet $\Gamma$. This notation is extended to sets of sequences as usual.
A \emph{trace} of $A$ is the projection $\rho | \Sigma$ of an execution $\rho$ of $A$. 
The set of executions, resp., traces, of an LTS $A$ is denoted by $E(A)$, resp., $Tr(A)$.
An LTS is \emph{deterministic} if for any state $s$ and any sequence $@t\in \Sigma^*$, there is at most
one state $s'$ such that $s\xrightarrow{@t}s'$.

@@

\noindent
We fix an arbitrary set $\<Ops>$ of operation identifiers, and for given sets
$\<Methods>$ and $\<Vals>$ of methods and values, we fix the sets
\begin{align*}
  & C = \set{ inv(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }
  \text{, and } \\
  & R = \set{ ret(m,d,k) : m \in \<Methods>, d \in \<Vals>, k \in \<Ops> }  
\end{align*}
of \emph{call actions} and \emph{return actions}; each call action $inv(m,d,k)$
combines a method $m \in \<Methods>$ and value $d \in \<Vals>$ with an
\emph{operation identifier} $k \in \<Ops>$. Operation identifiers are used to
pair call and return actions. We denote the operation identifier of a
call/return action $a$ by $\<op>(a)$. Call and return actions $c \in C$ and $r
\in R$ are \emph{matching}, written $c \match r$, when $\<op>(c) = \<op>(r)$. 
We may omit the second field from a call/return action $a$ for methods that have no inputs \textcolor{red}{(e.g., the pop method of a stack)}or return values \textcolor{red}{(e.g., the push method of a stack)} 
(e.g., the pop method of a stack).
A word $\tau \in @S^*$ over alphabet $@S$, such that $(C \u R) \subseteq @S$, is
\emph{well formed} when:
\begin{itemize}

@@

\textcolor{red}{For the below paragraph, I cannot see the equivalence  of informal explanation and formal definition of weakening. I think the formal one should be if a ret comes before a call in $h_1'$ then this order is preserved in $h_2$. But the definition says iff. Ignore this if I am wrong.}
Since libraries only dictate methods’ executions between their respective calls and returns, for any history they admit, they must also 
admit histories with weaker inter-operation ordering, in which calls may happen earlier, and/or returns later. This weakening relation
between histories can be formalized as follows: a history $h_1$ is \emph{weaker} than a history $h_2$, written $h_1 \sqsubseteq h_2$, 
if{f} there exists a well-formed history $h_1'$
obtained from $h_1$ by appending return actions, and deleting call actions,
such that:
\begin{quote}

@@

\end{quote}

$\overline{H}$ denotes the closure $\set{h: \exists h' \in H.\ h \sqsubseteq h'}$ of a 
history set $H$ under weakening.
A direct consequence of Definition~\ref{def:libraries} is that $H(L)=\overline{H(L)}$ for
every library $L$. A \emph{kernel} of L is any set $H$ such that $\overline{H} = L$ \textcolor{red}{(I think the correct expression is $\overline{H} = H(L)$ which makes the example set $H$ and the operator $H$  that extracts history from a library confusing with each other)}.
A library $L$ is called atomic if it has a sequential kernel. 
Atomic libraries are often considered as specifications for concurrent objects. 
In practice, libraries can be made atomic by guarding their methods bodies with global lock acquisitions.

@@

\textcolor{red}{A general organizational idea for  this subsection: Linearizability is defined in terms of weakening in which observational actions are just call and return events. Then, we put forward the idea of enriching $\Gamma$ by adding new actions. In the current organization, linearizability is defined after refinement before which this $\Gamma$ extension argument is explained. Hence, people may confuse that notion of linearizability is based on $\Gamma$ set we picked. I propose to first introduce linearizability. Then explain refinement and give Theorem 1.}
Conformance of a library $L_1$ to a specification given as an ``abstract'' library $L_2$ 
is formally captured by \emph{(observational) refinement}. Informally, given two libraries
$L_1$ and $L_2$ implementing the methods of some concurrent object, we
say $L_1$ refines $L_2$ if{f} every computation of every program
using $L_1$ would also be possible were $L_2$ used instead. As shown in~\citet{journals/tcs/FilipovicORY10,DBLP:conf/popl/BouajjaniEEH15},
refinement is equivalent to the \emph{linearizability} criterion~\cite{journals/toplas/HerlihyW90} when $L_2$ is an atomic library. 
Refinement can be proved using \emph{forward/backward simulation relations} which roughly, guarantee that steps
in the concrete library correspond to (sequences of) steps in the abstract library.

\begin{dfn}
\textcolor{red}{We assign same alphabet to both libraries in the definition but we say libraries can be defined over distinct alphabets as long as they share $\Gamma$ which is the case in our week machine. Why don't we define refinement like that at the first place?}
Let $L_1$ and $L_2$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$~\footnote{The two libraries could be defined over different alphabets as long as they share $\Gamma$.}. 
%agreeing on $\mathcal{M}$ and $\mathcal{D}$ sets. We define the set $A\Sigma = \{inv(m,d,i)| m \in \mathcal{M} \wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\} \cup \{ret(m,d,i)| m \in \mathcal{M}\wedge d \in \mathcal{D} \wedge i \in \mathbb{N}\}$ as abstract transition labels. Note that $A\Sigma \subseteq \Sigma_{L_1}$ and $A\Sigma \subseteq \Sigma_{L_2}$. 
The library $L_1$ \emph{$\Gamma$-refines} $L_2$ if{f} $Tr(L_1) | \Gamma \subseteq Tr(L_2) | \Gamma$.
%We say that $L_1$ \emph{refines} $L_2$ when $H(L_1) \subseteq H(L_2)$.
%for every trace $e \in Tr(L_1)$, there exists a trace $e' \in Tr(L_2)$ such that $e|A\Sigma$ is equivalent to $e'|A\Sigma$.
\end{dfn}
The actions in $\Gamma$ are called \emph{observable actions}. Usually, $\Gamma$ consists only of call and return actions \textcolor{red}{(For this case we say just refinement instead of $\Gamma-refinement$. We need to add this sentence since we say just refinement in Theorem 1.)}. As we show in the next sections, increasing the set of observable actions is useful in simplifying proofs of refinement.
%Linearizability is also a relation between two libraries and it is stricter than refinement. It requires $e'$ in Definition 2 to be a sequential one. A trace $e$ is sequential iff following two conditions hold for its projection to abstract transition labels $e|A\Sigma = e_1, ...e_n$: $(i)$ $e_1 = inv(m,d,k)$ for some $m \in \mathcal{M}, d \in \mathcal{D} and k \in \mathbb{N}$ and $(ii)$ for all $i \in [1,n)$, either $e_i = inv(m,d,k)$ and $e_{i+1} = ret(m,d',k)$ or $e_i = ret(m,d,k)$ and $e_{i+1} = inv(m',d',k')$ for some $m,m' \in \mathcal{M}$, $d,d' \in \mathcal{D}$ and $k,k' \in \mathbb{N}$.

@@

\textcolor{red}{An organizational concern: Do we really need the second definition of linearizability? I think our aim here is to introduce both refinement and linearizability and relate them via Thm 1. I cannot see any contribution of this part until Thm 1. You may ignore this message if I am missing something important about the second definition.}
The weakening order w.r.t. sequential histories can be also defined  in terms of ~\emph{linearization points}. 
%A linearization point 
%corresponds to a transition in that
%execution that occurs between the call and the return action of the operation.
A \emph{linearization point annotation} (LP-annotation for short) of a history $h$ is a sequence $\hat{h}$ obtained from $h$ by inserting actions $lin(m,d,k)$ with $m\in\<Methods>$, $d\in\<Vals>$ and $k\in \<Ops>$ such that:
\begin{itemize}
	\item every linearization point corresponds to an operation that has been called, i.e., TODO
	\item every linearization point occurs between the call and the return of the corresponding operation, i.e., TODO
	\item the return value of an operation is fixed by the linearization point, i.e., TODO
\end{itemize}
An LP-annotation is called \emph{complete} when every completed operation has a linearization point, i.e., TODO.
An operation $k$ is called \emph{effectuated} in a complete LP-annotation $\hat{h}$ when  $\hat{h}$ contains $lin(\_,\_,k)$.

@@

We define a notion of \emph{forward} simulation whose existence implies refinement, and it is equivalent to refinement when the right-hand side library $L_2$ is deterministic \textcolor{red}{(do we need to define deterministic?)}. A dual notion of \emph{backward} simulation is defined in Appendix~\ref{app:backSim}. The existence of backward simulations is equivalent to refinement for arbitrary libraries.
%We define a notion of forward simulation that is a sound approximation of refinement and a notion of backward simulation that precisely characterizes refinement. 
For a relation $R\subseteq A\times B$, $R[X]$ denotes the set of elements of $B$ related to elements of $X$ by $R$, i.e., $R[X]=\set{y:\exists x\in X.\ R(x,y)}$.

@@

\begin{dfn}
Let $L_1=(Q_1,\Sigma, q_0^1, \delta_1)$ and $L_2=(Q_2,\Sigma, q_0^2, \delta_2)$ be two libraries over a common alphabet $\Sigma$, and $\Gamma\subseteq \Sigma$ a set of actions such that $(C\cup R)\subseteq \Gamma$. A relation $\mathit{fs} \subseteq Q_{1} \times Q_{2}$ is called a \emph{$\Gamma$-forward simulation} from $L_1$ to $L_2$ iff the following holds:
\begin{itemize}
\item[(i)] $\mathit{fs}[s_0^1] = \{s_0^2 \}$ \textcolor{red}{Either we need to change $s_0^1$ to $q_0^1$ here or vice versa in the first line of the definition.}
\item[(ii-a)] If $(s,c,s') \in \delta_1$, for some $c\in C$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_0=c$, and $@s_i\in \Sigma\setminus(C\cup R)$ \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0<i<|@s|$.
%$@s = a_1, a_2, ..., a_n$ such that $a_1 = inv(m,d,k)$ and for all $i \in [2,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. The expression $u \xrightarrow{a} u'$ means that there exists a sequence of states $u_1, u_2,...,u_{n+1}$ such that $u_1 = u$, $u_{n+1} = u'$ and for all $i \in [1,n]$, $(u_i, a_i, u_{i+1}) \in \delta_{L_2}$.
\item[(ii-b)] If $(s,r,s') \in \delta_{1}$, for some $r\in R$, and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{@s} u'$, $@s_{|@s| -1}=r$, and $@s_i\in \Sigma\setminus(C\cup R)$  \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0\leq i<|@s| -1$.
% where $a = a_1, a_2, ..., a_n$ such that $a_n = ret(m,d,k)$ and for all $i \in [1,n-1]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. 
\item[(ii-c)] If $(s, \gamma , s') \in \delta_1$, for some $\gamma\in \Gamma\setminus (C\cup R)$, and $u \in fs[s]$, then there exists $u' \in fs[s']$ such that $u\xrightarrow{\gamma} u'  \in \delta_2$. \textcolor{red}{$(u,\gamma, u') \in \delta_2$ instead of $u\xrightarrow{\gamma} u'  \in \delta_2$?)}
\item[(ii-d)] If $(s,e,s') \in \delta_1$, for some $e \in \Sigma\setminus \Gamma$ and $u \in \mathit{fs}[s]$, then there exists $u' \in \mathit{fs}[s']$ such that $u \xrightarrow{\sigma} u'$ and $@s_i\in \Sigma\setminus(C\cup R)$  \textcolor{red}{($@s_i\in \Sigma\setminus \Gamma$)}, for each $0\leq i<|@s|$.  \textcolor{red}{(All of the above changes also apply to the backward simulation definition in appendix.)}
%where $a = a_1, a_2, ..., a_n$ such that for all $i \in [1,n]$, $a_i \in \Sigma_{L_2} \backslash A\Sigma $. Moreover, $a$ could be the empty sequence.
\end{itemize}
\end{dfn}
%If $\mathit{fs}[s]$ is a unique state for all $s \in Q_1$ then $\mathit{fs}$ is called a refinement mapping/function. 
A $\Gamma$-forward simulation establishes that $L_2$ can mimic every step of $L_1$ using a (possibly empty) sequence of steps. In order to imply $\Gamma$-refinement, steps of $L_1$ labeled by observable actions are treated in a particular way: a step of $L_1$ labeled by a call, resp., return, action is simulated by a sequence of steps of $L_2$ that start, resp., end, with the same action, and a step of $L_1$ labeled by another observable action should be matched by a step of $L_2$ labeled by the same action. The rest of the transitions in $L_1$ must be simulated by a possibly empty sequence of transitions of $L_2$ (with arbitrary labels).
